---
title: "R Notebook"
output: html_notebook
---

```{r}
library(beeswarm)
library(naniar)
```


# Make some utility functions

```{r}
load_file <- function(file_path){
  read.csv(file_path)
}



```


# Business Understanding

Need to add some things here about covid.

# Data Understanding

```{r}
# Load some of the data 
tx_data <- load_file("./../data/COVID-19_cases_TX.csv")
global_mobility_report <- load_file("./../data/Global_Mobility_Report.csv")
cases_plus_census <- load_file("./../data/COVID-19_cases_plus_census.csv")
```

## What data is availiable?
There is a large amount of data provided to us for this project.  We are given three different csv files:

* COVID-19_cases_plus_census
  * brief desc.
* COVID-19_cases_TX
  * brief desc.
* Global_Mobility_Report
  * brief desc.

All of these files contain important features; however, due to the large amount of data given we will focus on a few of the variables (features) that we believe are critically important.  First, lets narrow our focus down to the following variables: ____


## Verify Data Quality

### Is there missing data?
```{r}
cols_with_missing <- function(frame){
  colnames(frame)[ apply(frame, 2, anyNA) ]
}
```


```{r}
tx_missing <- tx_data[cols_with_missing(tx_data)]
global_mobility_report_missing <- global_mobility_report[cols_with_missing(global_mobility_report)]
cases_plus_census_missing <- cases_plus_census[cols_with_missing(cases_plus_census)]
```

```{r}
class(tx_data)
```



```{r}
# vis_miss(tx_missing, sort_miss = T)
# vis_miss(global_mobility_report, warn_large_data = F)
vis_miss(cases_plus_census_missing, sort_miss = T)

```




Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.


