---
title: "R Notebook"
output: html_notebook
---

```{r}
library(beeswarm)
library(naniar)
# install.packages("janitor")
library(janitor)
library(dplyr)
# install.packages("GGally")
library(sets)
library(tidyverse)
library(ggplot2)
library(GGally) # for ggpairs
# install.packages("maps")
library(maps)
# devtools::install_github(“UrbanInstitute/urbnmapr”)
```


# Make some utility functions

```{r}
load_file <- function(file_path){
  read.csv(file_path)
}



```


# Business Understanding

Need to add some things here about covid.

# Data Understanding

```{r}
# Load some of the data 
tx_data <- load_file("./../data/COVID-19_cases_TX.csv")
global_mobility_report <- load_file("./../data/Global_Mobility_Report.csv")
cases_plus_census <- load_file("./../data/COVID-19_cases_plus_census.csv")
```

## What data is availiable?
There is a large amount of data provided to us for this project.  We are given three different csv files:

* COVID-19_cases_plus_census
  * brief desc.
* COVID-19_cases_TX
  * brief desc.
* Global_Mobility_Report
  * brief desc.

All of these files contain important features; however, due to the large amount of data given we will focus on a few of the variables (features) that we believe are critically important.  First, lets narrow our focus down to the following variables: ____
```{r}
head(global_mobility_report)
```


## Verify Data Quality
```{r}
cols_keep <- c("county_fips_code", "confirmed_cases", "deaths", "median_income", "male_pop", "female_pop", "total_pop", "median_age", "worked_at_home")
# print(cols_keep)
subset_census <- cases_plus_census[cols_keep]
head(subset_census)
```
```{r}
sapply(subset_census, min, na.rm = TRUE)
```
```{r}
cols_keep <- c("date", "retail_and_recreation_percent_change_from_baseline", "grocery_and_pharmacy_percent_change_from_baseline", "parks_percent_change_from_baseline", "transit_stations_percent_change_from_baseline", "workplaces_percent_change_from_baseline", "residential_percent_change_from_baseline")
# print(cols_keep)
subset_mobility <- global_mobility_report[cols_keep]
```


```{r}
head(subset_mobility)
```
```{r}
# subset_mobility[['date'] <- as.Date(subset_mobility['date']], format='%m/%d/%y')
# head(subset_mobility)
subset_mobility$date <- as.Date(subset_mobility$date, format="%Y-%m-%d")
```



```{r}
# subset_mobility["date"] <- apply(subset_mobility["date"], 2, as.Date, format = '%m/%d/%y')
head(subset_mobility)

sapply(subset_mobility, min, na.rm = TRUE)
```
```{r}
summarise(min = min(subset_mobility, na.rm = TRUE),
            max = max(subset_mobility, na.rm = TRUE))
```
```{r}
print(min(subset_mobility$date, na.rm=TRUE))
print(max(subset_mobility$date, na.rm=TRUE))
```


### Is there missing data?
```{r}
cols_with_missing <- function(frame){
  colnames(frame)[ apply(frame, 2, anyNA) ]
}
```


```{r}
tx_missing <- tx_data[cols_with_missing(tx_data)]
global_mobility_report_missing <- global_mobility_report[cols_with_missing(global_mobility_report)]
cases_plus_census_missing <- cases_plus_census[cols_with_missing(cases_plus_census)]
```

```{r}
class(tx_data)
```



```{r}
# vis_miss(tx_missing, sort_miss = T)
# vis_miss(global_mobility_report, warn_large_data = F)
vis_miss(cases_plus_census_missing, sort_miss = T)

```

## Clean up DataFrames
First, we noticed that much of our data within the COVID-19_cases_TX file had many rows of data that are statewide and not just for a specific county.  Let's modify the data frame to be texas_county_info and texas_statewide_info.

```{r}
tx_county_info = filter(tx_data, county_name != "Statewide Unallocated")
tx_statewide_info = filter(tx_data, county_name == "Statewide Unallocated")
```


```{r}
head(tx_county_info)
```


```{r}
head(tx_statewide_info)
```

### Remove Columns from Data Frames

Remove columns from the dataframes that do not have any data, have the same data repeated (not useful since implicit knowledge)

```{r}
# Track the columns to see if we are removing any 
mobility_cols_orig <- colnames(global_mobility_report)
census_cols_orig <- colnames(cases_plus_census)
tx_county_cols_orig <- colnames(tx_county_info)
tx_statewide_cols_orig <- colnames(tx_statewide_info)

# Remove any columns that do not contain any data
global_mobility_report <- remove_empty(global_mobility_report)
cases_plus_census <- remove_empty(cases_plus_census)
tx_county_info <- remove_empty(tx_county_info)
tx_statewide_info <- remove_empty(tx_statewide_info)

# Remove columns that just have a constant value 
global_mobility_report <- remove_constant(global_mobility_report)
cases_plus_census <- remove_constant(cases_plus_census)
tx_county_info <- remove_constant(tx_county_info)
tx_statewide_info <- remove_constant(tx_statewide_info)

mobility_cols_new <- colnames(global_mobility_report)
census_cols_new <- colnames(cases_plus_census)
tx_county_cols_new <- colnames(tx_county_info)
tx_statewide_cols_new <- colnames(tx_statewide_info)

```

```{r}
# Get sets for columns 
print(as.set(mobility_cols_orig) - as.set(mobility_cols_new))
print(as.set(census_cols_orig) - as.set(census_cols_new))
print(as.set(tx_county_cols_orig) - as.set(tx_county_cols_new))
print(as.set(tx_county_cols_orig) - as.set(tx_statewide_cols_new))
```

When removing columns that either do not have data or a constant is repeated throughout the column, the global_mobility_report remains the same; however, the following columns are dropped for the respective dataframes:

* census
  * date, do_date, pop_15_and_over, pop_5_years_over, pop_divorced, pop_never_married, pop_now_married, pop_separated, pop_widowed, speak_only_english_at_home, speak_spanish_at_home,and speak_spanish_at_home_low_english
* tx_county_info
  * state, state_fips_code
* tx_statewide_info
  * county_fips_code, county_name, state, and state_fips_code
  
```{r}
head(tx_county_info)
```

### Are there Duplicates?
```{r}
mobility_duplicates <- duplicated(global_mobility_report)
census_duplicates <- duplicated(cases_plus_census)
tx_county_duplicates <- duplicated(tx_county_info)
tx_statewide_duplicates <- duplicated(tx_statewide_info)

print(sum(mobility_duplicates, na.rm = TRUE))
print(sum(census_duplicates, na.rm = TRUE))
print(sum(tx_county_duplicates, na.rm = TRUE))
print(sum(tx_statewide_duplicates, na.rm = TRUE))

```

We can see that there were no duplicate observations in the data.  At least none on the basis that the data was the exact same.

```{r}
unique(mobility_duplicates)
```

### Create columns based on population

```{r}
# make a pct infected column (virus so should get 1 time, but this is disputed for covid)
subset_census['pct_infected'] <- subset_census['confirmed_cases']/subset_census['total_pop']
subset_census['pct_deaths'] <- subset_census['deaths']/subset_census['total_pop']
head(subset_census)
```

```{r}
library(RColorBrewer)
plot_vs_county <- function(df, col_val, percentile=FALSE,
                           fips_title="county_fips_code", banks=6, 
                           legend_title="", graphic_title=""){
  # Subset for speed 
  df <- df[c(fips_title, col_val)]
  
  # Get county data
  gcounty <- ggplot2::map_data("county")
  # USA map data
  gusa <- map_data("state")
  
  if (banks > 9){
    mycolors <- colorRampPalette(brewer.pal(9, "Reds"))(banks)
  }
  
  # Format with subregions
  fipstab <-
      transmute(maps::county.fips, fips, county = sub(":.*", "", polyname)) %>%
      unique() %>%
      separate(county, c("region", "subregion"), sep = ",")
  
  # Combine in desired order (NA for missing)
  gcounty <- left_join(gcounty, fipstab, c("region", "subregion"))


  dis <- df
  dis$rprop <- rank(df[col_val])
  dis$pcls <- cut(100 * percent_rank(df[col_val]), seq(0, 100, len = banks),
                        include.lowest = TRUE)

  # Missing data
  anti_join(gcounty, dis, by = c("fips" = fips_title)) %>%
    select(region, subregion) %>%
    unique()
  gcounty_pop <- left_join(gcounty, dis, by = c("fips" = fips_title))
  fill_vals <- gcounty_pop[col_val]

  # Plot
  if (legend_title == ""){
    legend_title <- col_val
  }

  if (percentile == FALSE){
    # names(gcounty_pop)[names(gcounty_pop) == col_val] <- "col_of_interest"
    plt <- ggplot(gcounty_pop) +
      geom_polygon(aes(long, lat, group = group, fill = get(col_val)),
                   color = "grey", size = 0.1, name="Percent Infected") +
      geom_polygon(aes(long, lat, group = group),
                   fill = NA, data = gusa, color = "lightgrey") +
      coord_map("bonne", parameters = 41.6) + ggthemes::theme_map()+
      scale_fill_gradient2()
       # scale_fill_gradient(low = "white", high = "red", na.value = "grey")
      # scale_fill_gradientn(colours = terrain.colors(10))
  }

  if (percentile == TRUE){
    plt <- ggplot(gcounty_pop) +
      geom_polygon(aes(long, lat, group = group, fill = pcls),
                   color = "grey", size = 0.1) +
      geom_polygon(aes(long, lat, group = group),
                   fill = NA, data = gusa, color = "lightgrey") +
      coord_map("bonne", parameters = 41.6) + ggthemes::theme_map() +
      scale_fill_manual(values = mycolors, na.value = "grey") +
      # scale_fill_brewer(palette = "viridis", na.value = "grey") +
      theme(plot.title = element_text(family = "Helvetica", face = "bold", size = (15)),
            legend.background = element_rect(fill = NA), 
            legend.position = "left")
  }
  plt <- plt + labs(fill=legend_title) + ggtitle(graphic_title)
  plt
}
```


```{r}
plot_vs_county(subset_census, "pct_infected", legend_title = "Percent Infected")
plot_vs_county(subset_census, "pct_infected", percentile = TRUE, banks=11, 
               legend_title = "Percentile Infected",
               graphic_title = "Percentile of Percentage of People Infected by County")
```

```{r}
library(maps)

census_infected <- subset_census[c("county_fips_code", "pct_infected")]
# census_infected$color_density <- rainbow(n = 50, census_infected$pct_infected / max(census_infected$pct_infected))
census_infected$color_density <- heat.colors(50, census_infected$pct_infected/max(census_infected$pct_infected))
head(census_infected)
# county_infected_data <- 
# map("county", fill=TRUE, col=counties$color)
```

```{r}
data(county.fips)
## Set up fake df_pop_county data frame
df_pop_county <- data.frame(region=county.fips$fips)
df_pop_county$value <- county.fips$fips
y <- df_pop_county$value
df_pop_county$color <- gray(y / max(y))

## merge population data with county.fips to make sure color column is
## ordered correctly.
counties <- county.fips %>% left_join(df_pop_county, by=c('fips'='region'))
# print(head(counties))

newdata <- census_infected[order(census_infected$county_fips_code),]
newdata

# Through away counties that are not in default
new_data <- newdata[(newdata$county_fips_code %in% counties$fips),]
new_data
```


```{r}
data(county.fips)
## Set up fake df_pop_county data frame
df_pop_county <- data.frame(region=county.fips$fips)
df_pop_county$value <- county.fips$fips
y <- df_pop_county$value
df_pop_county$color <- gray(y / max(y))

## merge population data with county.fips to make sure color column is
## ordered correctly.
counties <- county.fips %>% left_join(df_pop_county, by=c('fips'='region'))
map("county", fill=TRUE, col=counties$color)
```
```{r}
census_infected
```

```{r}
data(county.fips)
df_pop_county <- data.frame(region=county.fips$fips)
df_pop_county
```

```{r}
counties <- county.fips %>% left_join(census_infected, by=c('fips'='county_fips_code'))
counties
```




```{r}
ggplot2::map_data("county")
as_tibble(maps::county.fips)
# dall <- ggplot2::map_data("county") %>% left_join(as_tibble(maps::county.fips))
```
```{r}
install.packages("mapproj")
```

```{r}
# http://homepage.stat.uiowa.edu/~luke/classes/STAT4580-2020/maps.html
# Choropleth Maps of County Population
library(ggplot2)
# install.packages(mapproj)
library(mapproj)

gcounty <- ggplot2::map_data("county")
ggplot(gcounty) +
    geom_polygon(aes(long, lat, group = group),
                 fill = NA, color = "black", size = 0.05) +
    coord_map("bonne", parameters = 41.6)
```


```{r}
head(filter(maps::county.fips, grepl(":", polyname)))
fipstab <-
    transmute(maps::county.fips, fips, county = sub(":.*", "", polyname)) %>%
    unique() %>%
    separate(county, c("region", "subregion"), sep = ",")
head(fipstab)
gcounty <- left_join(gcounty, fipstab, c("region", "subregion"))
head(gcounty)
```
```{r}
test <- gcounty %>% left_join(census_infected, by=c('fips'='county_fips_code'))
test
```
```{r}
install.packages("ggthemes")
library(ggthemes)
```
```{r}
census_infected
```
```{r}
census_infected %>% rank(pct_infected)
```


```{r}
ncls <- 6
w <- census_infected %>% select(fips = county_fips_code, pct_infected) %>% mutate(rpop = rank(pct_infected),
           pcls = cut(100 * percent_rank(pct_infected), seq(0, 100, len = ncls),
                      include.lowest = TRUE))
w
```
```{r}
anti_join(gcounty, w, "fips") %>%
    select(region, subregion) %>%
    unique()

gcounty_pop <- left_join(gcounty, w, "fips")
# filter(gcounty_pop, is.na(rpop)) %>%
#     select(region, subregion, pop, rpop, pcls) %>%
#     unique()
```
```{r}
gcounty_pop
```

```{r}
gusa <- map_data("state")
ggplot(gcounty_pop) +
    geom_polygon(aes(long, lat, group = group, fill = pct_infected),
                 color = "grey", size = 0.1, name="Percent Infected") +
    geom_polygon(aes(long, lat, group = group),
                 fill = NA, data = gusa, color = "lightgrey") +
    coord_map("bonne", parameters = 41.6) + ggthemes::theme_map()
```

```{r}
ggplot(gcounty_pop) +
    geom_polygon(aes(long, lat, group = group, fill = pcls),
                 color = "grey", size = 0.1) +
    geom_polygon(aes(long, lat, group = group),
                 fill = NA, data = gusa, color = "lightgrey") +
    coord_map("bonne", parameters = 41.6) + ggthemes::theme_map() +
    scale_fill_brewer(palette = "Reds", na.value = "blue",
                      name = "Percentile") +
    theme(legend.background = element_rect(fill = NA))
```
```{r}
plot_vs_county <- function(df, col_val, percentile=FALSE, fips_title="county_fips_code", banks=6, legend_title=""){
  # Subset for speed 
  df <- df[c(fips_title, col_val)]
  
  # Get county data
  gcounty <- ggplot2::map_data("county")
  # USA map data
  gusa <- map_data("state")
  
  # Format with subregions
  fipstab <-
      transmute(maps::county.fips, fips, county = sub(":.*", "", polyname)) %>%
      unique() %>%
      separate(county, c("region", "subregion"), sep = ",")
  
  # Combine in desired order (NA for missing)
  gcounty <- left_join(gcounty, fipstab, c("region", "subregion"))


  dis <- df
  dis$rprop <- rank(df[col_val])
  dis$pcls <- cut(100 * percent_rank(df[col_val]), seq(0, 100, len = banks),
                        include.lowest = TRUE)

  # Missing data
  anti_join(gcounty, dis, by = c("fips" = fips_title)) %>%
    select(region, subregion) %>%
    unique()
  gcounty_pop <- left_join(gcounty, dis, by = c("fips" = fips_title))
  fill_vals <- gcounty_pop[col_val]

  # Plot
  if (legend_title == ""){
    legend_title <- col_val
  }

  if (percentile == FALSE){
    # names(gcounty_pop)[names(gcounty_pop) == col_val] <- "col_of_interest"
    plt <- ggplot(gcounty_pop) +
      geom_polygon(aes(long, lat, group = group, fill = get(col_val)),
                   color = "grey", size = 0.1, name="Percent Infected") +
      geom_polygon(aes(long, lat, group = group),
                   fill = NA, data = gusa, color = "lightgrey") +
      coord_map("bonne", parameters = 41.6) + ggthemes::theme_map()
  }

  if (percentile == TRUE){
    plt <- ggplot(gcounty_pop) +
      geom_polygon(aes(long, lat, group = group, fill = pcls),
                   color = "grey", size = 0.1) +
      geom_polygon(aes(long, lat, group = group),
                   fill = NA, data = gusa, color = "lightgrey") +
      coord_map("bonne", parameters = 41.6) + ggthemes::theme_map() +
      scale_fill_brewer(palette = "Reds", na.value = "blue") +
      theme(legend.background = element_rect(fill = NA))
  }
  plt <- plt + labs(fill=legend_title)
  plt
}
```

```{r}
subset_census
# plot_vs_county(subset_census, "pct_infected", legend_title = "Pecent Infected")
```
```{r}
plot_vs_county(subset_census, "pct_infected", percentile = TRUE, legend_title = "Percentile Infected", banks = 10)
```



